\section{Operations and Maintenance}

\subsection{Operational Objectives}
Operations in AgriApp focus on keeping an edge node usable over long sessions with minimal friction. The practical objectives are straightforward: keep the service reachable, keep capture reliable, preserve storage headroom, and maintain recoverability when network conditions change.

The platform is therefore operated as a control loop rather than a static deployment: observe status, execute actions, verify outcomes, and log relevant context.

\subsection{Daily Runtime Workflow}
A typical operating cycle includes:
\begin{enumerate}
  \item verify backend health and storage,
  \item run one-shot validation if needed,
  \item start/stop unattended capture loop according to campaign plan,
  \item monitor gallery growth and metadata quality,
  \item apply cleanup/export actions before storage pressure builds up.
\end{enumerate}

This workflow is available via shell, browser, and Android app, which reduces dependency on any single interaction channel.

\subsection{Service Operations}
The backend server is managed by \texttt{systemctl --user}. In routine maintenance, the most relevant operations are status inspection, restart, and log tailing. API-level restart/reboot/poweroff actions provide remote control equivalents for field usage when shell access is inconvenient.

Operational safety remains important: reboot and poweroff are destructive from an availability perspective and should be used with explicit user confirmation and contextual logging.

On the Android side, these operations are grouped in the \texttt{System} tab using separate cards (\texttt{Storage}, \texttt{WiFi}, \texttt{Server}, \texttt{Power}). This reduces operator error by preventing accidental confusion between informational state (free storage), network-switch actions, and destructive power controls.

\subsection{Capture Operations}
Capture management now has two complementary modes:
\begin{itemize}
  \item one-shot capture for immediate validation (RPi or ESP source),
  \item unattended loop capture controlled by \texttt{/api/v1/capture/loop/*} (RPi) and \texttt{/api/v1/capture/esp/loop/*} (ESP).
\end{itemize}

For unattended operation, interval selection is an operational parameter (e.g., 30s and minute-scale presets in mobile UI or explicit API payloads). Start/stop semantics are explicit; operators should always verify loop status after transitions. In the Android control app, operators can start loops for \texttt{RPi}, \texttt{ESP}, or \texttt{Both}, while status remains source-specific.
Current ESP unattended operation is implemented through the same ESP proxy capture API used for one-shot, so metadata persistence behavior remains aligned between manual and automatic captures.

\subsection{Metadata Operations (GPS/BME280)}
Metadata is persisted opportunistically. Operations should treat missing sensor values as a diagnosable condition, not necessarily as a capture failure. If GPS or BME data disappears unexpectedly, the recommended sequence is:
\begin{enumerate}
  \item verify physical wiring/serial visibility,
  \item stop backend service if running manual GPS serial tests, to avoid dual readers on the same UART,
  \item run dedicated sensor test scripts,
  \item run one-shot capture and inspect generated metadata sidecar,
  \item only then escalate to service/device restart.
\end{enumerate}

This avoids unnecessary full restarts when the issue is local to one sensor path.
When ESP captures arrive without contextual values, operations now rely on metadata sidecar normalization/fallback at backend level to keep gallery behavior coherent while sensor diagnostics are performed.
At runtime, GPS access is additionally protected by a backend lock plus short cache TTL, which lowers serial contention during high-frequency health polling.
Operationally, the baseline setup currently keeps GT-U7 as active GPS module for data collection sessions, while NEO-6M remains an optional alternative for future retesting.
For ESP camera operation, field runs currently keep OV3660 as primary module; OV5640 remains out of baseline due to image artifact patterns observed during comparative tests.

\subsection{Storage Hygiene}
Because persistence is file-based, dataset hygiene is an explicit operational task. The platform supports single delete, delete-all, and dataset export. Maintainers should define a cadence for cleanup or off-device archival to avoid emergency low-space events.

Thumbnail cache is managed automatically by deletion endpoints, but periodic sanity checks remain useful in long-running campaigns.

\subsection{Network Mode Operations}
AgriApp is operated across heterogeneous network contexts. In infrastructure Wi-Fi mode, discovery should be fast and mostly cache-driven. In AP mode, local control continuity takes priority over throughput.

For AP reliability with ESP32-CAM devices, practical field configuration matters. In current operations, the rescue AP is managed with a fixed 2.4\,GHz channel (channel 6) and compatibility-oriented WPA2 settings. This avoids recurrent association failures observed when AP channel selection is left fully automatic or when security negotiation is too strict for older ESP firmware stacks.

Mode switching must be followed by a brief verification phase (reachability, status endpoint, shot action) before proceeding with high-value operations. This small discipline prevents silent mode drift during field sessions.
In current tablet operation, a delayed automatic discovery pass is triggered after Wi-Fi/AP mode changes, reducing the need for repeated manual discovery taps during subnet transitions.

Operational UI now exposes AP/Wi-Fi ON/OFF badges in the \texttt{Connections} tab and ``last seen'' timestamps for both RPi and ESP endpoints. This gives fast visual feedback about connectivity recency without adding verbose controls.

For Wi-Fi profile handling, operators can select a known profile and trigger \texttt{Connect now} directly from the \texttt{System/WiFi} card. This action maps to backend network-mode switching with preferred profile selection and is intended for controlled transitions between preconfigured environments.

\subsection{Incident Response}
A practical incident playbook is:
\begin{enumerate}
  \item check physical power/connectivity,
  \item check \texttt{/api/v1/health}, \texttt{/api/v1/health/checklist}, and \texttt{/api/v1/system/status},
  \item check capture path (RPi first, then ESP proxy),
  \item check sensor visibility if metadata is missing,
  \item restart service, then reboot only if necessary.
\end{enumerate}

This order minimizes downtime and preserves diagnostic signal.

\subsection{Operational Logging}
Effective troubleshooting depends on correlated logs. The project uses a combination of backend logs, systemd status/logs, and Android in-app logs. Operators should record at least three contextual tags when reporting incidents: active network mode, active host/IP, and last successful action.
Recent app iterations keep in-app logging structured but concise (discovery milestones, network switch milestones, loop lifecycle changes, and shot outcomes). Detailed contextual values are expected to be inspected from metadata artifacts and gallery views rather than verbose control-log lines.
On the web side, runtime logging is exposed through a single unified stream, avoiding split views between server and ESP sections and reducing operator confusion during active troubleshooting.

\subsection{Maintenance Outlook}
Maintenance effort should continue reducing legacy complexity. The most valuable ongoing work is progressive legacy endpoint retirement, tighter contract tests on \texttt{/api/v1}, and continued UI/ops simplification so that field operation stays predictable even as features grow.
